\begin{thebibliography}{99}

\bibitem{Alstadsaeter2019}
Alstads{\ae}ter, A., Johannesen, N., \& Zucman, G. (2019). Tax evasion and inequality. \textit{American Economic Review, 109}(6), 2073–2103. 

\bibitem{Alstadsaeter2022}
Alstadsæter, A., Zucman, G., Planterose, B. \& Økland, A. (2022) Who owns offshore real estate? Evidence from Dubai (EU Tax Observatory working paper 1). https://www.bankofengland.co.uk/-/media/boe/files/events/2022/june/workshop-hf-and-alstadsaeter-paper-2.pdf [Accessed 14th Feburary 2024].

\bibitem{Alstadsaeter2024}
Alstads{\ae}ter, A., Johannesen, N., Le Guern Herry, S., \& Zucman, G. (2024). \textit{Global Tax Evasion Report 2024}. EU Tax Observatory.

\bibitem{Allingham1972}
Allingham, M. G., \& Sandmo, A. (1972). Income tax evasion: A theoretical analysis. \textit{Journal of Public Economics, 1}(3-4), 323–338. 

\bibitem{Angrist2008}
Angrist, J. D., \& Pischke, J.-S. (2008). Mostly harmless econometrics. Princeton University Press.

\bibitem{Blondel2008}
Blondel, V. D., Guillaume, J.-L., Lambiotte, R., \& Lefebvre, E. (2008). Fast unfolding of communities in large networks. \textit{Journal of Statistical Mechanics: Theory and Experiment, 2008}(10), P10008. 

\bibitem{Bourdieu1977}
Bourdieu, P. (1977). \textit{Outline of a theory of practice} (R. Nice, Trans.). Cambridge University Press.

\bibitem{Burke1774}
"Speech of Edmund Burke, Esq. on American taxation, April 19, 1774:." In the digital collection Eighteenth Century Collections Online. https://name.umdl.umich.edu/004902755.0001.000. University of Michigan Library Digital Collections. Accessed May 27, 2025.

\bibitem{Bustos2022}
Bustos, S., Gidi, V., \& Pomeranz, D. (2022). \textit{The race between tax enforcement and tax planning: Evidence from a natural experiment in Chile} (Working Paper).

\bibitem{Carter2017}
Carter, P. (2017). \textit{Why do Development Finance Institutions use offshore financial centres?} Overseas Development Institute.

\bibitem{Chang2023a}
Ho-Chun Herbert Chang, Brooke Harrington, Feng Fu, Daniel N Rockmore, Complex systems of secrecy: the offshore networks of oligarchs, PNAS Nexus, Volume 2, Issue 3, March 2023, pgad051

\bibitem{Chang2023b}
Ho-Chun Herbert Chang, Brooke Harrington, Feng Fu, Daniel N Rockmore (2023). Secrecy Strategies: Global Patterns in Elites' Quest for Confidentiality in Offshore Finance. Presented at The Central Bank of The Bahamas AML/CFT Conference. \url{https://bahamasamlconference.centralbankbahamas.com/documents/2024-03-26-15-06-53-Session-2---Secrecy-Strategies-Global-Patterns-in-Elites-Quest-for-Confidentiality-in-Offshore-Finance.pdf}

\bibitem{Chetty2014}
Chetty, R., Hendren, N., Kline, P., \& Saez, E. (2014). Where is the land of opportunity? The geography of intergenerational mobility in the United States. \textit{The Quarterly Journal of Economics, 129}(4), 1553–1623.

\bibitem{Hearson2020}
Christensen, R.C. \& Hearson, M. (2019) The new politics of global tax governance: taking stock a decade after the financial crisis. Review of International Political Economy, 26(5), 1068–1088. 

\bibitem{Christensen2020}
Christensen, R.C. (2020) Elite professionals in transnational tax governance. Global Networks, 21, 265–293. 

\bibitem{Christensen2024}
Christensen, R. C. (2024). Harnessing network power: Weaponised interdependence in global tax policy. Global Policy.

\bibitem{Christensen2021}
Christensen, R.C., Seabrooke, L. \& Wigan, D. (2021) Professional action in global wealth chains. Regulation \& Governance, 16, 721. 

\bibitem{DeGroen2017}
De Groen, W. P. (2017). Role of advisors and intermediaries in the schemes revealed in the Panama Papers. Directorate-General for Internal Policies. Study for the Pana Committee

\bibitem{Emmenegger2015}
Emmenegger, P. (2015) The long arm of justice: U.S. structural power and international banking. Business \& Politics, 17(3), 473–493. 

\bibitem{Farrell2023}
Farrell, H. \& Newman, A. (2023b) Underground empire: how America weaponized the world economy. London: Allen Lane.

\bibitem{FATF2023}
Financial Action Task Force. (2025). \textit{International standards on combating money laundering and the financing of terrorism \& proliferation: The FATF Recommendations}. FATF. Retrieved from \url{https://www.fatf-gafi.org/en/publications/Fatfrecommendations/Fatf-recommendations.html}

\bibitem{GarciaBernardo2017}
Garcia-Bernardo, J., Fichtner, J., \& Takes, F. W. (2017). Uncovering offshore financial centers: Conduits and sinks in the global corporate ownership network. \textit{Scientific Reports, 7}(1), 6246. 

\bibitem{Granovetter1973}
Granovetter, M. S. (1973). The strength of weak ties. \textit{American Journal of Sociology, 78}(6), 1360–1380. 

\bibitem{Harrington2016}
Harrington, B. (2016). \textit{Capital without borders: Wealth managers and the one percent}. Harvard University Press.

\bibitem{Hastie2009}
Hastie, T., Tibshirani, R., \& Friedman, J. (2009). \textit{The elements of statistical learning: Data mining, inference, and prediction} (2nd ed.). Springer.

\bibitem{Ho2009}
Ho, K. C. (2009). Liquidated. An Ethnography of Wall Street. Duke University Press.

\bibitem{Hoang2022}
Hoang, K. (2022). \textit{Spiderweb Capitalism: How Global Elites Exploit Frontier Markets}. Princeton University Press.

\bibitem{Hoang2020}
Hoang, Kimberly Kay (2020). Engendering Global Capital: How Homoerotic Triangles Facilitate Foreign Investments into Risky Markets. Gender and Society 34 (4):547-572.

\bibitem{ICIJ}
International Consortium of Investigative Journalists. (n.d.). \textit{Offshore Leaks Database}. Retrieved from \url{https://offshoreleaks.icij.org/}

\bibitem{Kejriwal2020}
Kejriwal, S., \& Dang, V. A. (2020). Unveiling the hidden global network of financial secrecy: A panama papers perspective. \textit{EPJ Data Science, 9}(1), 25. 

\bibitem{Kleven2011}
Kleven, H. J., Knudsen, M. B., Kreiner, C. T., Pedersen, S., \& Saez, E. (2011). Unwilling or unable to cheat? Evidence from a tax audit experiment in Denmark. \textit{Econometrica, 79}(3), 651–692. 

\bibitem{Laffitte2024}
Laffitte, S. (2024). \textit{The market for tax havens} [Data set and working paper]. Paris School of Economics.

\bibitem{LondonoVelez2021}
Londo{\~n}o-V{\'e}lez, J., \& {\'A}vila-Mahecha, J. (2021). Enforcing wealth taxes in the developing world: Quasi-experimental evidence from Colombia. \textit{American Economic Review: Insights, 3}(2), 131–148. 

\bibitem{Lowendahl2005}
L{\o}wendahl, B. R. (2005). \textit{Strategic management of professional service firms} (3rd ed.). Copenhagen Business School Press.

\bibitem{Maister1993}
Maister, D. H. (1993). \textit{Managing the professional service firm}. Free Press.

\bibitem{Neely2022}
Neely, M. H. (2022). \textit{Hedged Out: Inequality and Insecurity on Wall Street}. University of California Press.

\bibitem{Newman2010}
Newman, M. E. J. (2010). \textit{Networks: An introduction}. Oxford University Press.

\bibitem{OECD2014}
Organisation for Economic Co-operation and Development. (2014). \textit{Standard for automatic exchange of financial account information in tax matters}. OECD Publishing. 

\bibitem{Palan2002}
Palan, R. (2002). Tax havens and the commercialization of state sovereignty. \textit{International Organization, 56}(1), 151–176. 

\bibitem{Palan2010}
Palan, R., Murphy, R., \& Chavagneux, C. (2010). \textit{Tax havens: How globalization really works}. Cornell University Press.

\bibitem{Piketty2014}
Piketty, T. (2014). \textit{Capital in the twenty-first century} (A. Goldhammer, Trans.). Harvard University Press.

\bibitem{Rixen2010}
Rixen, T. (2010). From double tax avoidance to tax competition: Explaining the institutional trajectory of international tax governance. Review of International Political Economy, 18(2), 197–227. 

\bibitem{Russell1946}
Russell, B. (1946). \textit{History of Western philosophy}. George Allen \& Unwin.

\bibitem{Saez2019}
Saez, E., \& Zucman, G. (2019). \textit{The triumph of injustice: How the rich dodge taxes and how to make them pay}. W. W. Norton \& Company.

\bibitem{Seabrooke2017}
Seabrooke, L., \& Wigan, D. (2017). The governance of global wealth chains. \textit{Review of International Political Economy, 24}(1), 1-29. 

\bibitem{Shiller2012}
Shiller, R. J. (2012). \textit{Finance and the good society}. Princeton University Press.

\bibitem{Slemrod2019}
Slemrod, J. (2019). Tax Compliance and Enforcement: New Developments. \textit{Journal of Economic Literature, 57}, 907-954. 

\bibitem{Stausholm2024}
Stausholm, S. \& Garcia-Bernardo, J. (2024) Unfollow the money: mapping the micro agents of international tax. Review of International Political Economy, 31(4), 1197–1219. 

\bibitem{Stiglitz2025}
Stiglitz, J. E. (2025, April). America Is becoming the World's Largest Tax Haven \textit{Project Syndicate}. Retrieved from \url{https://www.project-syndicate.org/commentary/america-becoming-largest-tax-haven-under-trump-by-joseph-e-stiglitz-2025-04}

\bibitem{TJN2022}
Tax Justice Network. (2022). Financial Secrecy Index 2022. Tax Justice Network. https://fsi.taxjustice.net/

\bibitem{Tilly1990}
Tilly, C. (1990). \textit{Coercion, capital, and European states, AD 990-1990}. Basil Blackwell.

\bibitem{Torslov2018}
T{\o}rsl{\o}v, T. R., Wier, L. S., \& Zucman, G. (2018). \textit{The missing profits of nations} (NBER Working Paper No. 24701). National Bureau of Economic Research. 

\bibitem{Vdem2023}
Varieties of Democracy (V-Dem) Institute. (2023). \textit{V-Dem Dataset v13}. Retrieved from \url{https://www.v-dem.net/data/the-v-dem-dataset/}

\bibitem{vonNordenflycht2010}
von Nordenflycht, A. (2010). What is a professional service firm? Toward a theory and taxonomy of knowledge-intensive firms. \textit{Academy of Management Review, 35}(1), 155–174. \url{https://doi.org/10.5465/amr.35.1.zok155}

\bibitem{Wei2024}
Wei, X. (2024). \textit{The development of Hong Kong as an intermediate financial centre after 1978, by studying Chinese state-owned banks in Hong Kong} (Unpublished doctoral thesis). City, University of London.

\bibitem{Weinberg2023}
ICIJ. \textit{Cyprus confidential}. International Consortium of Investigative Journalists. Retrieved from \url{https://www.icij.org/investigations/cyprus-confidential/}

\bibitem{Zar2010}
Zar, J.H. (2010) Biostatistical Analysis. 5th Edition, Prentice-Hall/Pearson

\bibitem{Zucman2015}
Zucman, G. (2015) The hidden wealth of nations: the scourge of tax havens. Chicago: University of Chicago Press.

\end{thebibliography}

